Active Learning for SAT-Solver Benchmarking

### 2022-09-27 paper/paper.tex
Replace \textit with \emph
... for highlighting.

\begin{abstract}
  Experimentation is one of the main phases in algorithm engineering, including the propositional-sat\-is\-fia\-bi\-li\-ty problem (SAT).
  While the field of SAT-solver benchmarking is well established, traditional bench\-mark-set-selection approaches do not optimize for benchmark runtime.
  Benchmark-set selection is the process of choosing a subset of instances out of a pool to be representative and reliable in discriminating SAT solvers.
  This paper presents an active-learning-based benchmark-set-selec\-tion approach that solves the \textit{New-Solver Problem}:
  Given a new solver, we want to find its ranking amidst its competitors with as little runtime as possible while still giving accurate ranking predictions.
  We can determine a new solver's rank regarding the SAT Competition 2022 Anniversary Track instances in about \SI{10}{\%} of the time it would take to run the solver on all instances of this dataset.
  We achieve this while still maintaining a prediction accuracy of about \SI{92}{\%} regarding the PAR-2 ranking of this solver if it were evaluated on the complete dataset.
  Additionally, we discuss the importance of instance families in the selection process. 
  Our tool provides the engineer of propositional-satisfiability solvers with a faster way of determining a new SAT-solver's performance in comparison to sampling the benchmark set by hand.

@@

This current article is proposing a novel active-learning-based benchmark-set-selection approach.
Our algorithm is unique in that it is the only one, to the best of our knowledge, that has all of the following features (also refer to Tab.~\ref{tab:requirements}).
Our benchmarking process outputs a \textit{ranking} to compare the new solver against the existing set of solvers as described above.
We optimize the \textit{runtime} that our strategy needs to arrive at its conclusion.
We use problem-instance and runtime \textit{features} $\varphi\!: \mathcal{A} \times \mathcal{I} \rightarrow \mathbb{R}^n$ since such features already have proven to be helpful in classic benchmark-set-selection approaches~\cite{HoosKSS13} as well as in portfolio selection~\cite{CollauttiMMO13,NgokoCT19}.
%JB: I think we can remove this criterion, as it also is satisified by all competitors (and is commented out in table; if we keep criterion, we should also show it in table)
Moreover, we select instances \textit{non-randomly} and \textit{incrementally}.
By doing so, we have a greater impact on the properties of the benchmarking approach such as its required runtime.
Sampling instances incrementally means taking runtime information from already done experiments into account when choosing the next.
Rather than solely outputting a binary classification, i.e, the new solver is worse than solver A, we provide a \textit{scoring function} that shows by which margin a solver is worse and how similar it is to existing approaches (\textit{interpretability}).
%JB: we should rename "interpretability" criterion, term can have lots of different meanings and our specific meaning is not very intuitive
Finally, our strategy's cost \textit{scales linearly} with the cardinality of the set of existing solvers $\mathcal{A}$ to compare against and the instances $\mathcal{I}$ within our pool of possible benchmark instances.
We avoid comparing pairs of solvers.
This would scale quadratically.
%JB: need to discuss this; log-linear effort might be sufficient if binary-search-/recursive-sorting-like procedure used

@@

\section{Related Work}
This section discusses work in several related fields.
First, we take a look at traditional SAT-solver benchmarking.
Second, we explore how SAT-solver configuration systems deal with navigating possible configurations.
Third, we discuss an existing benchmark-set-selection approach that optimizes for benchmark runtime.
Finally, we show how active-learning techniques can help to tackle the aforementioned \textit{New-Solver Problem}.

@@

Hoos et. al.~\cite{HoosKSS13} discuss which properties are most desirable when selecting a set of SAT benchmark instances, for instance for a competition.
The selection strategy presented is static, i.e., it does not depend on the solvers to distinguish.
Desirable properties are a variety of instances (avoids over-fitting of solver approaches), adapted instance hardness (not too easy but also not too hard; solvers cannot be distinguished if too many time-out or finish immediately), and no duplicates (having duplicates gives those instances more weight).
To avoid having too similar instances, they make use of a distance-based approach using the SATzilla features~\cite{XuHHL08,features}.
Those features comprise general instance-size features and graph-representation features among others.
%JB: we should probably cut the discussion of each approach's criteria to save space, maybe only mention the unsatisifed ones
Their approach \textit{ranks} solvers by providing its PAR-2 score, is \textit{feature-based}, and is \textit{scalable} in the sense that the evaluation time scales linearly with more instances and solvers.
Also, their approach is \textit{interpretable} in the sense that the obtained PAR-2 score induces a distance between solvers.
However, they do not optimize for benchmark \textit{runtime} and select instances \textit{randomly}, apart from constraints on the instance hardness and feature distance.

@@

While this approach considers the configuration's \textit{runtime}, is \textit{feature-based}, and \textit{scales linearly} by having a fixed time budget, it is not capable of the other three features listed in Tab.~\ref{tab:requirements}.
First, an algorithm configurator cannot be used to \textit{rank} a new solver since it only looks at promising solvers/configurations rather than the overall average performance.
If the new solver does not belong to this category (new ideas are seldomly the best approach right from the start, i.e., they need some tweaking), we do not have any information on how this solver performed.
%JB: "this category" = category of most promising solvers? not 100% clear
Second, while using a model-based selection strategy to sample configurations, instance sampling is done \textit{randomly}, i.e., without building a model over instances.
And third, their approach is not \textit{interpretable} regarding the new solver performance since we are already missing the feature to \textit{rank} the new solver among existing ones.
%JB: well, the internal meta-model of the configurator allows to estimate the solver performance for particular configurations (instead of comparing solvers), which allows a specific performance assessment, so the approach is interpretable in that direction (though for comparison to other solvers, only top performance should be interesting)

@@

Regarding our desired features of adaptive benchmark-set-selection (Tab.~\ref{tab:requirements}), the aforementioned approach already ticks off the first four features, i.e., it \textit{ranks} solvers, optimizes for \textit{runtime}, is \textit{feature-based}, and uses \textit{incremental non-random} sampling.
Although they provide a confidence metric for their prediction, which makes this approach partially \textit{interpretable}, it is not clear how similar two given solvers are or on which instances they behave similarly, as the prediction is solely binary.
Moreover, the major shortcoming lies within it being not \textit{scalable} with the number solvers.
As they compare only pairs of solvers, scaling with more and more solvers and ideas becomes an issue as there is a quadratic amount of pairs to distinguish.
%JB: maybe rather log-linear than quadratic, see a prior comment
Since a new solver idea is seldomly best at first try but requires more or less tweaking, it is desired to rank a new solver in a way that scales linearly with the number of existing approaches.
%JB: this general motivation should go there where linearity is introduced as criterion, not here
In this way, a new solver idea can be discarded if it performs poorly across the board or may be further tweaked if it shows promising results at least in some cases.

\subsubsection{Active learning (AL).}
%JB: section does not make clear which of the following approaches we will use (reads more like fundamentals than related work)
The posed \textit{New-Solver Problem} has stark similarities to the well-studied field of active learning within recommender systems, especially the \textit{new user problem}~\cite{RubensESK15}.
The classic exploration-exploitation dilemma arises since, on the one hand, we want to maximize the utility an instance provides to our model and, on the other hand, minimize the cost (CPU time) that is associated with its acquisition.
%JB: no sure it is arises here as well; dilemma is typical for searching one optimal solution in continuous space; however, if we find an interesting/useful instance here, I'm not sure if we would continue picking similar instances ("exploting" instead of "exploring") just because we expect them to be useful as well

@@

AL algorithms can be categorized into \textit{synthesis-bas\-ed} \cite{0001AEMN22,GarzonMG22,2019gaal} and \textit{pool-bas\-ed} approaches \cite{GolbandiKL11,distribAL,HarpaleY08,KapoorGUD07,KorenBV09,KornerW06,MelvilleM04,SinhaED19,TongK01}.
%JB: we can probably cut some citations, the three most important should be sufficient
While synthesis-based methods (Fig. \ref{fig:activesynth}) generate instances that might be interesting to label, pool-based methods (Fig. \ref{fig:activepool}) rely on a pool of unlabeled instances from which to draw items.

@@

Pool-based methods can be further refined into \textit{Error-based} \cite{GolbandiKL11,KorenBV09}, \textit{Un\-cer\-tain\-ty-based} \cite{HarpaleY08,KapoorGUD07,TongK01}, \textit{Distribution-based} methods \cite{distribAL,SinhaED19} and \textit{Ensembles} \cite{KornerW06,MelvilleM04}.
Error-based strategies pick instances that have the most influence on the model parameters.
Uncertainty-based methods sample instances close to their perceived decision boundary.
Distribution-based approaches sample instances whose features are unaccounted for amongst the already sampled instances.
Finally, ensembles are a meta-strategy that combines the predictions of several different active-learning models into a selection choice.

\subsubsection{Summary.}
To sum up this section, we have shown both traditional and recent benchmark-set-selection approaches.
Moreover, we have briefly outlined how active-learning techniques can help to incrementally select benchmark sets.
In the following sections, we present an active-learning-based benchmark-set-selection approach that solves the \textit{New-Solver Problem} while optimizing for runtime and being both interpretable and scalable.

\section{Active Learning for SAT-Solver Benchmarking}
\label{sec:main}
Our algorithm follows a simple three-step schema (Alg.~\ref{algALBenchmark}).
%JB: as it is rather simple and has different parts to be plugged in, I think we should rebrand it as "framework" rater than algorithm
After initializing the model $\mathcal{M}$, we incrementally \textit{select} a problem instance from the pool of instances, one at a time, evaluate the new solver $\hat{a}$ on it, and \textit{update} the model with the acquired result.
We \textit{stop} when meeting some stopping condition.
Since the potential runtime of experiments is by magnitudes larger than the model's update time, we only consider incrementing our benchmark set by one instance at a time rather than using batches.
% For now, we also restrict ourselves to the aforementioned sequential one-instance-at-a-time process.
At the core of the model $\mathcal{M}$ is a prediction function $f\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \mathcal{V}$ that powers the decisions within the previously mentioned three steps.
Note that the prediction target values $\mathcal{V}$ are unspecified to allow for more flexibility within our particular approaches.
Refer to the next sections for more details.

@@

We use a function $\gamma\!: {\mathcal{A} \times \mathcal{I}} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$ to transform runtime observations.
In our experiments, this function is a clustering algorithm.
Note that this transformation is done per instance to adapt to different instance difficulties. 
Also, we define ${\gamma\!\left(a, i\right) := k}$ iff $r\!\left(a, i\right) = \,\perp$ for all $a \in \mathcal{A}$ and $i \in \mathcal{I}$.
To determine a ranking based on those labels, we define
\begin{equation}
  \operatorname{score}\!\left(a\right) := \frac{1}{|\mathcal{I}|} \sum_{i \in \mathcal{I}} w_{a,i} \cdot \left(\gamma\!\left(a, i\right) - 1\right) \enspace \textrm{,}
  %JB: does the "-1" provide any benefits, or does it only shift scores?
  \label{eq:rankingeq}
\end{equation}
with $w_{a,i} := 2$ (PAR-\textit{2} penalty factor) if $\gamma\!\left(a, i\right) = k$ else $w_{a,i} := 1$.
To be applicable to $\hat{a}$, we extend $\operatorname{score}$ by replacing $\gamma\!\left(a, i\right)$ by the model's predictions $f\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$.
%JB: as "$\operatorname{score}$" is not visible, we should use some other way of highlighting

@@

That discretized runtimes are sufficient for distinguishing solvers can also be shown by comparing the fraction of solver pairs whose runtimes are significantly different.
Within the SAT~Competition~2022 Anniversary Track, about \SI{89.95}{\%} of solver pairs are significantly different (according to a Wilcoxon-signed-rank test with $\alpha = 0.05$).
While only time-out labels (solver times out or not on a particular instance) are sufficient to significantly distinguish about \SI{87.04}{\%} of solver pairs, discrete-runtime labels are sufficient to distinguish about \SI{87.83}{\%} of solver pairs.
%\footnote{Based on hierarchical clustering with a log-single-link criterion for $k = 3$.}
This means that we do not lose much discriminatory power using discrete runtime labels.
The number of significant comparisons between solvers \textit{alone}, though, does only offer limited insights as we might achieve significant differences in all pairs by assigning some randomly skewed values as labels.
%JB: thus, we should reorder section to mention significant difference (this paragraph) first and deciding solver pairs (previous paragraph, stronger argument) second (rather than giving the stronger argument first, making the second argument superfluous)
However, the predictiveness of the PAR-2 score, the labels' discriminatory power, as well as that it enables us to use classification at the same time, show the clear advantage of runtime discretization.
%JB: at the same time with what? with the other benefits?

\subsection{Instance Sampling}
\label{sec:main2}
Our instance-sampling strategies (also refer to Alg.~\ref{algALBenchmark}; line 3) can be divided into two categories.
First, we present possible baseline approaches including existing work on how to pick \textit{interesting} instances~\cite{Bossek021a,GentHJKMNN14,MatriconAFSH21}.
%JB: approaches under "Second" are also at least "inspired by existing work" and probably also pick "interesting" instances, distinction between the two categories not entirely clear
Second, we show promising machine-learning-based sampling strategies that make use of the model's label-prediction function $f$ and are inspired by existing work within the realms of active learning~\cite{settles2009active}.

@@

Bossek and Wagner~\cite{Bossek021a} introduce a custom ranking-score metric incorporating the number of \textit{good} pairs and \textit{bad} pairs.
%JB: some selection methods, like this one, are not mentioned later any more (in particular, not in experimental design), so we don't need to describe them in detail (maybe mention shortly if used in preliminary experiments, but turned out to be not useful)
A good solver pair, regarding a particular instance, are solvers that perform according to their global ranking.
So if solver $a$ is better than solver $b$ on instance $i$ and $a$ is also in total better than $b$, it is a \textit{good} solver pair on $i$.
Otherwise, it is a \textit{bad} pair.
Based on this classification, they define a 3-tuple for each instance consisting of the number of good pairs, the accumulated performance differences of bad pairs, and the accumulated performance differences of good pairs.
By selecting instances with the highest score according to the lexicographical order of the tuples, they try to favor instances that are predictive of the final ranking.
This assumes, though, that the new solver $\hat{a}$ behaves similarly to the existing solvers.

@@

The Wilcoxon stopping criterion stops the iterative active-learning algorithm when we are on average certain enough that the predicted runtime labels are sufficiently different in comparison to the other solvers.
To be more specific, we stop when
\begin{equation}
  W_{\hat{a}} := \frac{1}{\left\lvert \mathcal{A} \right\rvert} \sum_{a \in \mathcal{A}} \operatorname{wilcoxon}_{\alpha}\!\left(\gamma\!\left(a, \mathcal{I}\right), f\!\left(\hat{a}, \mathcal{I}\right) \right)
\end{equation}
drops below a certain threshold.
%JB: if this is info gain regarding runtimes, doesn't Eq. (4) mean that we would prefer instances with low entropy of runtime distribution after considering the new solver, i.e., instances that are not very discriminatory?
Thereby, \textit{wilcoxon} denotes the $p$-value of a Wilcoxon signed-rank test of the runtime-label distributions and $\gamma\!\left(a, \mathcal{I}\right)$ and $f\!\left(\hat{a}, \mathcal{I}\right)$ are the expected set extensions on $\gamma$ and $f$ respectively.
%JB: "set extension" might be rather technical and hard to understand; also, why "expected"? in the sense of "straightforward"?
To improve the stability of this criterion, we use an exponential moving average
\begin{equation}
  W_{\exp}^{\left(m\right)} := \beta W_{\hat{a}} + \left(1 - \beta\right) W_{\exp}^{\left(m - 1\right)} \enspace \textrm{,}
\end{equation}
whereby $m$ denotes the count of iterations within Alg.~\ref{algALBenchmark} and $W_{\exp}^{\left(0\right)} := 1$.

@@

\subsection{Optimization Goal}
\label{sec:goal}
As already stated in the introductory section, this work addresses the \textit{New-Solver Problem} (cf.~Def.~\ref{def:new-solver-problem}).
By running Alg.~\ref{algALBenchmark}, we obtain a prediction model $\mathcal{M}$ that provides us with a scoring function \textit{score}.
Also, it outputs the fraction of runtime that is needed to arrive at its conclusion.
First and foremost, our goal is to provide the engineer of new SAT solvers with an accurate ranking.
We define the \textit{ranking accuracy} $O_{\operatorname{acc}} \in \left[0, 1\right]$ (higher is better) by the fraction of pairs $\left(\hat{a}, a\right)$ for all $a \in \mathcal{A}$ that are decided correctly by the given ranking.
For now, we exclude the equality of solvers since data shows that the PAR-2 scores of all our solvers are different.
So, possible ranking decisions may be, for example, solver $a$ is better than $b$ or $b$ is better than $a$.
The ranking accuracy is affected by whether this decision is correctly made.
Matricon et. al.~\cite{MatriconAFSH21} also evaluate their approach by the fraction of solver pairs that is correctly decided.
Second, we also have to optimize for runtime.
The \textit{fraction of runtime} that the algorithm needs to arrive at its conclusion is denoted by $O_{\operatorname{rt}} \in \left[0, 1\right]$ (lower is better).
This metric puts the runtime summed over the sampled instances in relation to the runtime summed over all instances in the dataset.
Overall, we want to find an approach that maximizes
\begin{equation}
  O_\delta := \delta O_{\operatorname{acc}} + \left(1 - \delta\right) \left(1 - O_{\operatorname{rt}}\right) \enspace \textrm{,}
  \label{eq:opt}
\end{equation} 
whereby $\delta \in \left[0, 1\right]$ allows for linear weighting between the two optimization goals $O_{\operatorname{acc}}$ and $O_{\operatorname{rt}}$.
Plotting the approaches that maximize $O_\delta$ for all $\delta \in \left[0, 1\right]$ on a $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram provides us with a Pareto front of the best approaches for different optimization-goal weightings.

>>>>

\begin{abstract}
  Experimentation is one of the main phases in algorithm engineering, including the propositional-sat\-is\-fia\-bi\-li\-ty problem (SAT).
  While the field of SAT-solver benchmarking is well established, traditional bench\-mark-set-selection approaches do not optimize for benchmark runtime.
  Benchmark-set selection is the process of choosing a subset of instances out of a pool to be representative and reliable in discriminating SAT solvers.
  This paper presents an active-learning-based benchmark-set-selec\-tion approach that solves the \emph{New-Solver Problem}:
  Given a new solver, we want to find its ranking amidst its competitors with as little runtime as possible while still giving accurate ranking predictions.
  We can determine a new solver's rank regarding the SAT Competition 2022 Anniversary Track instances in about \SI{10}{\%} of the time it would take to run the solver on all instances of this dataset.
  We achieve this while still maintaining a prediction accuracy of about \SI{92}{\%} regarding the PAR-2 ranking of this solver if it were evaluated on the complete dataset.
  Additionally, we discuss the importance of instance families in the selection process. 
  Our tool provides the engineer of propositional-satisfiability solvers with a faster way of determining a new SAT-solver's performance in comparison to sampling the benchmark set by hand.

@@

This current article is proposing a novel active-learning-based benchmark-set-selection approach.
Our algorithm is unique in that it is the only one, to the best of our knowledge, that has all of the following features (also refer to Tab.~\ref{tab:requirements}).
Our benchmarking process outputs a \emph{ranking} to compare the new solver against the existing set of solvers as described above.
We optimize the \emph{runtime} that our strategy needs to arrive at its conclusion.
We use problem-instance and runtime \emph{features} $\varphi\!: \mathcal{A} \times \mathcal{I} \rightarrow \mathbb{R}^n$ since such features already have proven to be helpful in classic benchmark-set-selection approaches~\cite{HoosKSS13} as well as in portfolio selection~\cite{CollauttiMMO13,NgokoCT19}.
%JB: I think we can remove this criterion, as it also is satisified by all competitors (and is commented out in table; if we keep criterion, we should also show it in table)
Moreover, we select instances \emph{non-randomly} and \emph{incrementally}.
By doing so, we have a greater impact on the properties of the benchmarking approach such as its required runtime.
Sampling instances incrementally means taking runtime information from already done experiments into account when choosing the next.
Rather than solely outputting a binary classification, i.e, the new solver is worse than solver A, we provide a \emph{scoring function} that shows by which margin a solver is worse and how similar it is to existing approaches (\emph{interpretability}).
%JB: we should rename "interpretability" criterion, term can have lots of different meanings and our specific meaning is not very intuitive
Finally, our strategy's cost \emph{scales linearly} with the cardinality of the set of existing solvers $\mathcal{A}$ to compare against and the instances $\mathcal{I}$ within our pool of possible benchmark instances.
We avoid comparing pairs of solvers.
This would scale quadratically.
%JB: need to discuss this; log-linear effort might be sufficient if binary-search-/recursive-sorting-like procedure used

@@

\section{Related Work}
This section discusses work in several related fields.
First, we take a look at traditional SAT-solver benchmarking.
Second, we explore how SAT-solver configuration systems deal with navigating possible configurations.
Third, we discuss an existing benchmark-set-selection approach that optimizes for benchmark runtime.
Finally, we show how active-learning techniques can help to tackle the aforementioned \emph{New-Solver Problem}.

@@

Hoos et. al.~\cite{HoosKSS13} discuss which properties are most desirable when selecting a set of SAT benchmark instances, for instance for a competition.
The selection strategy presented is static, i.e., it does not depend on the solvers to distinguish.
Desirable properties are a variety of instances (avoids over-fitting of solver approaches), adapted instance hardness (not too easy but also not too hard; solvers cannot be distinguished if too many time-out or finish immediately), and no duplicates (having duplicates gives those instances more weight).
To avoid having too similar instances, they make use of a distance-based approach using the SATzilla features~\cite{XuHHL08,features}.
Those features comprise general instance-size features and graph-representation features among others.
%JB: we should probably cut the discussion of each approach's criteria to save space, maybe only mention the unsatisifed ones
Their approach \emph{ranks} solvers by providing its PAR-2 score, is \emph{feature-based}, and is \emph{scalable} in the sense that the evaluation time scales linearly with more instances and solvers.
Also, their approach is \emph{interpretable} in the sense that the obtained PAR-2 score induces a distance between solvers.
However, they do not optimize for benchmark \emph{runtime} and select instances \emph{randomly}, apart from constraints on the instance hardness and feature distance.

@@

While this approach considers the configuration's \emph{runtime}, is \emph{feature-based}, and \emph{scales linearly} by having a fixed time budget, it is not capable of the other three features listed in Tab.~\ref{tab:requirements}.
First, an algorithm configurator cannot be used to \emph{rank} a new solver since it only looks at promising solvers/configurations rather than the overall average performance.
If the new solver does not belong to this category (new ideas are seldomly the best approach right from the start, i.e., they need some tweaking), we do not have any information on how this solver performed.
%JB: "this category" = category of most promising solvers? not 100% clear
Second, while using a model-based selection strategy to sample configurations, instance sampling is done \emph{randomly}, i.e., without building a model over instances.
And third, their approach is not \emph{interpretable} regarding the new solver performance since we are already missing the feature to \emph{rank} the new solver among existing ones.
%JB: well, the internal meta-model of the configurator allows to estimate the solver performance for particular configurations (instead of comparing solvers), which allows a specific performance assessment, so the approach is interpretable in that direction (though for comparison to other solvers, only top performance should be interesting)

@@

Regarding our desired features of adaptive benchmark-set-selection (Tab.~\ref{tab:requirements}), the aforementioned approach already ticks off the first four features, i.e., it \emph{ranks} solvers, optimizes for \emph{runtime}, is \emph{feature-based}, and uses \emph{incremental non-random} sampling.
Although they provide a confidence metric for their prediction, which makes this approach partially \emph{interpretable}, it is not clear how similar two given solvers are or on which instances they behave similarly, as the prediction is solely binary.
Moreover, the major shortcoming lies within it being not \emph{scalable} with the number solvers.
As they compare only pairs of solvers, scaling with more and more solvers and ideas becomes an issue as there is a quadratic amount of pairs to distinguish.
%JB: maybe rather log-linear than quadratic, see a prior comment
Since a new solver idea is seldomly best at first try but requires more or less tweaking, it is desired to rank a new solver in a way that scales linearly with the number of existing approaches.
%JB: this general motivation should go there where linearity is introduced as criterion, not here
In this way, a new solver idea can be discarded if it performs poorly across the board or may be further tweaked if it shows promising results at least in some cases.

\subsubsection{Active learning (AL).}
%JB: section does not make clear which of the following approaches we will use (reads more like fundamentals than related work)
The posed \emph{New-Solver Problem} has stark similarities to the well-studied field of active learning within recommender systems, especially the \emph{new-user problem}~\cite{RubensESK15}.
The classic exploration-exploitation dilemma arises since, on the one hand, we want to maximize the utility an instance provides to our model and, on the other hand, minimize the cost (CPU time) that is associated with its acquisition.
%JB: no sure it is arises here as well; dilemma is typical for searching one optimal solution in continuous space; however, if we find an interesting/useful instance here, I'm not sure if we would continue picking similar instances ("exploting" instead of "exploring") just because we expect them to be useful as well

@@

AL algorithms can be categorized into \emph{synthesis-bas\-ed} \cite{0001AEMN22,GarzonMG22,2019gaal} and \emph{pool-bas\-ed} approaches \cite{GolbandiKL11,distribAL,HarpaleY08,KapoorGUD07,KorenBV09,KornerW06,MelvilleM04,SinhaED19,TongK01}.
%JB: we can probably cut some citations, the three most important should be sufficient
While synthesis-based methods (Fig. \ref{fig:activesynth}) generate instances that might be interesting to label, pool-based methods (Fig. \ref{fig:activepool}) rely on a pool of unlabeled instances from which to draw items.

@@

Pool-based methods can be further refined into \emph{Error-based} \cite{GolbandiKL11,KorenBV09}, \emph{Un\-cer\-tain\-ty-based} \cite{HarpaleY08,KapoorGUD07,TongK01}, \emph{Distribution-based} methods \cite{distribAL,SinhaED19} and \emph{Ensembles} \cite{KornerW06,MelvilleM04}.
Error-based strategies pick instances that have the most influence on the model parameters.
Uncertainty-based methods sample instances close to their perceived decision boundary.
Distribution-based approaches sample instances whose features are unaccounted for amongst the already sampled instances.
Finally, ensembles are a meta-strategy that combines the predictions of several different active-learning models into a selection choice.

\subsubsection{Summary.}
To sum up this section, we have shown both traditional and recent benchmark-set-selection approaches.
Moreover, we have briefly outlined how active-learning techniques can help to incrementally select benchmark sets.
In the following sections, we present an active-learning-based benchmark-set-selection approach that solves the \emph{New-Solver Problem} while optimizing for runtime and being both interpretable and scalable.

\section{Active Learning for SAT-Solver Benchmarking}
\label{sec:main}
Our algorithm follows a simple three-step schema (Alg.~\ref{algALBenchmark}).
%JB: as it is rather simple and has different parts to be plugged in, I think we should rebrand it as "framework" rater than algorithm
After initializing the model $\mathcal{M}$, we incrementally \emph{select} a problem instance from the pool of instances, one at a time, evaluate the new solver $\hat{a}$ on it, and \emph{update} the model with the acquired result.
We \emph{stop} when meeting some stopping condition.
Since the potential runtime of experiments is by magnitudes larger than the model's update time, we only consider incrementing our benchmark set by one instance at a time rather than using batches.
% For now, we also restrict ourselves to the aforementioned sequential one-instance-at-a-time process.
At the core of the model $\mathcal{M}$ is a prediction function $f\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \mathcal{V}$ that powers the decisions within the previously mentioned three steps.
Note that the prediction target values $\mathcal{V}$ are unspecified to allow for more flexibility within our particular approaches.
Refer to the next sections for more details.

@@

We use a function $\gamma\!: {\mathcal{A} \times \mathcal{I}} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$ to transform runtime observations.
In our experiments, this function is a clustering algorithm.
Note that this transformation is done per instance to adapt to different instance difficulties. 
Also, we define ${\gamma\!\left(a, i\right) := k}$ iff $r\!\left(a, i\right) = \,\perp$ for all $a \in \mathcal{A}$ and $i \in \mathcal{I}$.
To determine a ranking based on those labels, we define
\begin{equation}
  \operatorname{score}\!\left(a\right) := \frac{1}{|\mathcal{I}|} \sum_{i \in \mathcal{I}} w_{a,i} \cdot \left(\gamma\!\left(a, i\right) - 1\right) \enspace \textrm{,}
  %JB: does the "-1" provide any benefits, or does it only shift scores?
  \label{eq:rankingeq}
\end{equation}
with $w_{a,i} := 2$ (PAR-\emph{2} penalty factor) if $\gamma\!\left(a, i\right) = k$ else $w_{a,i} := 1$.
To be applicable to $\hat{a}$, we extend $\operatorname{score}$ by replacing $\gamma\!\left(a, i\right)$ by the model's predictions $f\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$.
%JB: as "$\operatorname{score}$" is not visible, we should use some other way of highlighting

@@

That discretized runtimes are sufficient for distinguishing solvers can also be shown by comparing the fraction of solver pairs whose runtimes are significantly different.
Within the SAT~Competition~2022 Anniversary Track, about \SI{89.95}{\%} of solver pairs are significantly different (according to a Wilcoxon-signed-rank test with $\alpha = 0.05$).
While only time-out labels (solver times out or not on a particular instance) are sufficient to significantly distinguish about \SI{87.04}{\%} of solver pairs, discrete-runtime labels are sufficient to distinguish about \SI{87.83}{\%} of solver pairs.
%\footnote{Based on hierarchical clustering with a log-single-link criterion for $k = 3$.}
This means that we do not lose much discriminatory power using discrete runtime labels.
The number of significant comparisons between solvers \emph{alone}, though, does only offer limited insights as we might achieve significant differences in all pairs by assigning some randomly skewed values as labels.
%JB: thus, we should reorder section to mention significant difference (this paragraph) first and deciding solver pairs (previous paragraph, stronger argument) second (rather than giving the stronger argument first, making the second argument superfluous)
However, the predictiveness of the PAR-2 score, the labels' discriminatory power, as well as that it enables us to use classification at the same time, show the clear advantage of runtime discretization.
%JB: at the same time with what? with the other benefits?

\subsection{Instance Sampling}
\label{sec:main2}
Our instance-sampling strategies (also refer to Alg.~\ref{algALBenchmark}; line 3) can be divided into two categories.
First, we present possible baseline approaches including existing work on how to pick \emph{interesting} instances~\cite{Bossek021a,GentHJKMNN14,MatriconAFSH21}.
%JB: approaches under "Second" are also at least "inspired by existing work" and probably also pick "interesting" instances, distinction between the two categories not entirely clear
Second, we show promising machine-learning-based sampling strategies that make use of the model's label-prediction function $f$ and are inspired by existing work within the realms of active learning~\cite{settles2009active}.

@@

Bossek and Wagner~\cite{Bossek021a} introduce a custom ranking-score metric incorporating the number of \emph{good} pairs and \emph{bad} pairs.
%JB: some selection methods, like this one, are not mentioned later any more (in particular, not in experimental design), so we don't need to describe them in detail (maybe mention shortly if used in preliminary experiments, but turned out to be not useful)
A good solver pair, regarding a particular instance, are solvers that perform according to their global ranking.
So if solver $a$ is better than solver $b$ on instance $i$ and $a$ is also in total better than $b$, it is a \emph{good} solver pair on $i$.
Otherwise, it is a \emph{bad} pair.
Based on this classification, they define a 3-tuple for each instance consisting of the number of good pairs, the accumulated performance differences of bad pairs, and the accumulated performance differences of good pairs.
By selecting instances with the highest score according to the lexicographical order of the tuples, they try to favor instances that are predictive of the final ranking.
This assumes, though, that the new solver $\hat{a}$ behaves similarly to the existing solvers.

@@

The Wilcoxon stopping criterion stops the iterative active-learning algorithm when we are on average certain enough that the predicted runtime labels are sufficiently different in comparison to the other solvers.
To be more specific, we stop when
\begin{equation}
  W_{\hat{a}} := \frac{1}{\left\lvert \mathcal{A} \right\rvert} \sum_{a \in \mathcal{A}} \operatorname{wilcoxon}_{\alpha}\!\left(\gamma\!\left(a, \mathcal{I}\right), f\!\left(\hat{a}, \mathcal{I}\right) \right)
\end{equation}
drops below a certain threshold.
%JB: if this is info gain regarding runtimes, doesn't Eq. (4) mean that we would prefer instances with low entropy of runtime distribution after considering the new solver, i.e., instances that are not very discriminatory?
Thereby, \emph{wilcoxon} denotes the $p$-value of a Wilcoxon signed-rank test of the runtime-label distributions and $\gamma\!\left(a, \mathcal{I}\right)$ and $f\!\left(\hat{a}, \mathcal{I}\right)$ are the expected set extensions on $\gamma$ and $f$ respectively.
%JB: "set extension" might be rather technical and hard to understand; also, why "expected"? in the sense of "straightforward"?
To improve the stability of this criterion, we use an exponential moving average
\begin{equation}
  W_{\exp}^{\left(m\right)} := \beta W_{\hat{a}} + \left(1 - \beta\right) W_{\exp}^{\left(m - 1\right)} \enspace \textrm{,}
\end{equation}
whereby $m$ denotes the count of iterations within Alg.~\ref{algALBenchmark} and $W_{\exp}^{\left(0\right)} := 1$.

@@

\subsection{Optimization Goal}
\label{sec:goal}
As already stated in the introductory section, this work addresses the \emph{New-Solver Problem} (cf.~Def.~\ref{def:new-solver-problem}).
By running Alg.~\ref{algALBenchmark}, we obtain a prediction model $\mathcal{M}$ that provides us with a scoring function \emph{score}.
Also, it outputs the fraction of runtime that is needed to arrive at its conclusion.
First and foremost, our goal is to provide the engineer of new SAT solvers with an accurate ranking.
We define the \emph{ranking accuracy} $O_{\operatorname{acc}} \in \left[0, 1\right]$ (higher is better) by the fraction of pairs $\left(\hat{a}, a\right)$ for all $a \in \mathcal{A}$ that are decided correctly by the given ranking.
For now, we exclude the equality of solvers since data shows that the PAR-2 scores of all our solvers are different.
So, possible ranking decisions may be, for example, solver $a$ is better than $b$ or $b$ is better than $a$.
The ranking accuracy is affected by whether this decision is correctly made.
Matricon et. al.~\cite{MatriconAFSH21} also evaluate their approach by the fraction of solver pairs that is correctly decided.
Second, we also have to optimize for runtime.
The \emph{fraction of runtime} that the algorithm needs to arrive at its conclusion is denoted by $O_{\operatorname{rt}} \in \left[0, 1\right]$ (lower is better).
This metric puts the runtime summed over the sampled instances in relation to the runtime summed over all instances in the dataset.
Overall, we want to find an approach that maximizes
\begin{equation}
  O_\delta := \delta O_{\operatorname{acc}} + \left(1 - \delta\right) \left(1 - O_{\operatorname{rt}}\right) \enspace \textrm{,}
  \label{eq:opt}
\end{equation} 
whereby $\delta \in \left[0, 1\right]$ allows for linear weighting between the two optimization goals $O_{\operatorname{acc}}$ and $O_{\operatorname{rt}}$.
Plotting the approaches that maximize $O_\delta$ for all $\delta \in \left[0, 1\right]$ on a $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram provides us with a Pareto front of the best approaches for different optimization-goal weightings.
